### 2014-02-06 thesis-proposal/problem.tex
Updated missing references

%In a conventional \emph{monolithic} programming system, support for each of these features would need to be built into the language and tools. 
No system today builds in support for all of the features enumerated above. Instead, libraries generally provide support for regular expressions by leveraging  general-purpose abstraction mechanisms. Unfortunately, it is impossible to fully encode the syntax and the specialized static and dynamic semantics described above in terms of general-purpose notations and abstractions. Library providers thus need to compromise, typically by asking clients to provide regular expressions as strings and deferring parsing, typechecking and compilation to run-time. This introduces performance overhead and can lead to unanticipated run-time errors (as shown in \cite{spishak2012type}) and security vulnerabilities (due to injection attacks when user inputs are spliced into patterns, for example) \cite{string-stuff}. Useful tools for working with regular expressions are rarely integrated into editors, and even more rarely in a way that facilitates their discovery and use directly when the developer is manipulating regular expressions. Tools that must be discovered independently and accessed externally are used infrequently \cite{emersonStudy} and can be more awkward than necessary \cite{Campbell:2008:DRT:1636642.1636651, Omar:2012:ACC:2337223.2337324}, leading to lower productivity \cite{emersonStudy, Omar:2012:ACC:2337223.2337324}.\todo{complete references}

@@

When the semantics or implementation of a system must be extended to fully realize a new feature, as in the example above, providers typically take a \emph{language-external approach}, either by developing a new or derivative programming system (supported by \emph{language workbenches} \cite{erdweg2013state},  \emph{DSL frameworks} \cite{fowler2010domain} or \emph{compiler generators} \cite{abc}), or by extending an existing system by some mechanism that is not part of the language itself, such as an extension mechanism for a {particular} compiler\footnote{Compilers that modify, or allow modification of, the semantics of their base language, rather than simply permitting semantics-preserving optimizations, should be considered a pernicious means for creating new languages. Many ``practical'' compilers, including \texttt{gcc}, \texttt{GHC} and \texttt{SML/NJ}, are guilty of this, meaning that some programs that seem to be written in C, Haskell or Standard ML are actually written in tool-specific derivatives of these languages. Language-integrated mechanisms do not lead to such fragmentation.} or other tool. For example, a researcher interested in providing regular expression related features (let us refer to these collectively as \texttt{R}) might design a new system with built-in support for these, perhaps basing it on an existing system containing some general-purpose features (\texttt{G}). A different researcher developing a new language-based parallel programming abstraction or implementation strategy (\texttt{P}) might  take the same strategy. A third researcher, developing an alternative parallel programming abstraction (\texttt{Q}) might again do the same. This results in a collection of distinct systems as diagrammed in Figure \ref{approaches}a. Unfortunately, when providers of new features take language-external approaches like this, it causes  problems for clients related to \textbf{orthogonality} and \textbf{client compatibility}, described below. %This latter method couples the semantics of the feature to the implementation details of a particular tool. Because the use of one implementation entails a different semantics for the feature than another, the extended tool acts, \emph{de facto}, as a distinct system for our purposes.

>>>>

%In a conventional \emph{monolithic} programming system, support for each of these features would need to be built into the language and tools. 
No system today builds in support for all of the features enumerated above. Instead, libraries generally provide support for regular expressions by leveraging  general-purpose abstraction mechanisms. Unfortunately, it is impossible to fully encode the syntax and the specialized static and dynamic semantics described above in terms of general-purpose notations and abstractions. Library providers thus need to compromise, typically by asking clients to provide regular expressions as strings and deferring parsing, typechecking and compilation to run-time. This introduces performance overhead and can lead to unanticipated run-time errors (as shown in \cite{spishak2012type}) and security vulnerabilities (due to injection attacks when user inputs are spliced into patterns, for example) \cite{Bravenboer:2007:PIA:1289971.1289975}. Useful tools for working with regular expressions are rarely integrated into editors, and even more rarely in a way that facilitates their discovery and use directly when the developer is manipulating regular expressions. Tools that must be discovered independently and accessed externally are used infrequently \cite{Murphy-Hill:2011:PIE:1958824.1958888} and can be more awkward than necessary \cite{Campbell:2008:DRT:1636642.1636651, Omar:2012:ACC:2337223.2337324}, leading to lower productivity.

@@

When the semantics or implementation of a system must be extended to fully realize a new feature, as in the example above, providers typically take a \emph{language-external approach}, either by developing a new or derivative programming system (supported by \emph{language workbenches} \cite{erdweg2013state},  \emph{DSL frameworks} \cite{fowler2010domain} or \emph{compiler generators} \cite{brooker1963compiler}), or by extending an existing system by some mechanism that is not part of the language itself, such as an extension mechanism for a {particular} compiler\footnote{Compilers that modify, or allow modification of, the semantics of their base language, rather than simply permitting semantics-preserving optimizations, should be considered a pernicious means for creating new languages. Many ``practical'' compilers, including \texttt{gcc}, \texttt{GHC} and \texttt{SML/NJ}, are guilty of this, meaning that some programs that seem to be written in C, Haskell or Standard ML are actually written in tool-specific derivatives of these languages. Language-integrated mechanisms do not lead to such fragmentation.} or other tool. For example, a researcher interested in providing regular expression related features (let us refer to these collectively as \texttt{R}) might design a new system with built-in support for these, perhaps basing it on an existing system containing some general-purpose features (\texttt{G}). A different researcher developing a new language-based parallel programming abstraction or implementation strategy (\texttt{P}) might  take the same strategy. A third researcher, developing an alternative parallel programming abstraction (\texttt{Q}) might again do the same. This results in a collection of distinct systems as diagrammed in Figure \ref{approaches}a. Unfortunately, when providers of new features take language-external approaches like this, it causes  problems for clients related to \textbf{orthogonality} and \textbf{client compatibility}, described below. %This latter method couples the semantics of the feature to the implementation details of a particular tool. Because the use of one implementation entails a different semantics for the feature than another, the extended tool acts, \emph{de facto}, as a distinct system for our purposes.
